---
title: "December 3, 2018"
author: "Rachel Havranek"
date: "12/4/2018"
output: html_document
---
# Load Libraries and functions 
```{r import libraries and functions, message=FALSE}
#setwd("/Volumes/HAVRANEK18/Lab_testing_flasks")  
library(ggplot2)
library (plotly)
library (lubridate)
library(tidyverse)
library(grid)
library(hms)
library(fpp2)
library(zoo)

#derivative function 
deriv <- function(x, y) {
  diffvar <- diff(y) / diff(x)
}

peaks.read <- function(filepath){
  	files <-  list.files(path=filepath, recursive = T,full.names = T) 
	#lists the file names in the filepath directory
  	tables <- lapply(files, read.table,header = T, stringsAsFactors = F) 
	#reads in the files
  	df <- do.call("rbind",tables) #combines all the tables into one dataframe
  	df2 <- df[,c(1:2,17:19)] #subsets the dataframe to include only necessary columns
  	df3 <- df2[complete.cases(df2),] #subsets the dataframe to include only 
	#complete cases, eliminating rows at the end that aren't complete
  	return(df3)
	} #peaks.read is a function that will read files from all sub-folders of the
	#directory named in filepath and combine them into one dataframe

h2o_slope <- function(data,window){
  H2O_Slope_Window <- numeric()
  for (i in 1:nrow(data)){    
    if(i<=window/2){                                                                   
      H2O_Slope_Window[i] = 
        as.numeric(coef(lm(data$H2O[1:(i+window/2)]~
		data$seconds[1:(i+window/2)]))[2])
    }else{
      if(i >= (nrow(data)-window/2)){
        H2O_Slope_Window[i] = 
          as.numeric(coef(lm(data$H2O[(i-window/2):nrow(data)]~
		data$seconds[(i-window/2):nrow(data)]))[2])
      }else{
        H2O_Slope_Window[i] = 
          as.numeric(coef(lm(data$H2O[(i-window/2):(i+window/2)]~
		data$seconds[(i-window/2):(i+window/2)]))[2])       
      }
    }
  }
  return(H2O_Slope_Window)
} #h2o_slope will calculate the slope of h2o vs. experiment second in a moving window  
```

# Load Data 
```{r}
# folder_120318 <- "/Volumes/HAVRANEK18/120318"
# data_120318 <- peaks.read(folder_120318)
# saveRDS(data_120318, "/Volumes/HAVRANEK18/Lab_testing_flasks/data/data_120318.RDS")
data_120318 <- readRDS("/Volumes/HAVRANEK18/Lab_testing_flasks/data/data_120318.RDS")
```

# Add in some Metadata
This chunk (1) creates a date/time column to match my timezone, so I can use my lab notes as a guide for data reduction, (2) adds in a linearity correction column for later comparison, (3) parses the data down to when I was running experiments, and (4) plots that water concentration data 

Determined by altering push flow: 
\[\delta^{18}O_{corr} = \delta^{18}O_{raw}+ (-0.0002*[H2O]+4.8021)\]
\[\delta^{2}H = \delta^{2}H_{raw} + ( -0.0004*[H2O]+5.3133)\]
```{r}

data_120318 <- data_120318 %>% mutate(
  lin_d18O = (-0.0002*H2O+4.8021) + Delta_18_16,
    lin_dD = (-0.0004*H2O +5.31 ) + Delta_D_H,
  MST = 
    paste(data_120318$DATE, data_120318$TIME) %>% 
    ymd_hms() %>% 
    with_tz(tzone=c("America/Denver"))
)

data_120318 <- data_120318 %>% 
  filter(MST >  "2018-11-26 15:00:00")

# These labels help me tie to my notes more easily 

#These labels help me visually in my plots 
label1 <- grobTree(textGrob("Bottle 16", x=0.2,  y=0.95, hjust=0,
  gp=gpar(col="red", fontsize=13, fontface="italic")))

label2 <- grobTree(textGrob("Bottle 15", x=0.45,  y=0.95, hjust=0,
  gp=gpar(col="red", fontsize=13, fontface="italic")))

label3 <- grobTree(textGrob("Bottle 14", x=0.78,  y=0.95, hjust=0,
  gp=gpar(col="red", fontsize=13, fontface="italic")))

label4 <- grobTree(textGrob("Probe 2", x=0.05,  y=0.95, hjust=0,
  gp=gpar(col="red", fontsize=13, fontface="italic")))

label5 <- grobTree(textGrob("Probe 1", x=0.34,  y=0.95, hjust=0,
  gp=gpar(col="red", fontsize=13, fontface="italic")))

label6 <- grobTree(textGrob("Probe 1", x=0.57,  y=0.95, hjust=0,
  gp=gpar(col="red", fontsize=13, fontface="italic")))

label7 <- grobTree(textGrob("Probe 2", x=0.66,  y=0.98, hjust=0,
  gp=gpar(col="red", fontsize=13, fontface="italic")))

label8 <- grobTree(textGrob("Probe 1", x=0.9,  y=0.95, hjust=0,
  gp=gpar(col="red", fontsize=13, fontface="italic")))

all_120318 <- ggplot(data_120318, aes(MST, H2O))+
  labs(x="Time (MST)", title="November 26, 2018")+
  geom_point()+
  theme_classic()+annotation_custom(label1)+
  annotation_custom(label2)+
  annotation_custom(label3)+
  annotation_custom(label4)+
  annotation_custom(label5)+
  annotation_custom(label6)+
  annotation_custom(label7)+
  annotation_custom(label8)

all_120318
```

```{r}
Out_120318 <- data_120318 

Out_120318$seconds <- as.numeric(row.names(Out_120318))
 
Out_120318 <- Out_120318 %>% 
  mutate(
  m_11= rollmean(H2O, k=11, fill=NA),
  m_25=rollmean(H2O, k=25, fill=NA)) %>% 
  subset(seconds >13 & seconds<(nrow(Out_120318)-13))
```


## Calulate the first derivative and second derivative of water concentration data, then plot the second derivative 
```{r}
#Initialize data frame
Out_120318_fd<-as.data.frame(1:(nrow(Out_120318)-1))

# mutate in the first derivatives, and the rolling first derivatives 
Out_120318_fd<- Out_120318_fd %>% 
  mutate(
    fd = deriv(Out_120318$seconds, Out_120318$H2O),
    fd11=deriv(Out_120318$seconds, Out_120318$m_11),
    fd25=deriv(Out_120318$seconds, Out_120318$m_25)
  )

#Force column names so I can do the second derivative 
colnames(Out_120318_fd, do.NULL = FALSE)
colnames(Out_120318_fd)<-c("seconds", "fd", "fd11", "fd25")

#Initialize second derivative data frame
Out_120318_sd <- as.data.frame(1:(nrow(Out_120318_fd)-1))

#Mutate in the second derivatives 
Out_120318_sd <- Out_120318_sd %>% 
  mutate(
    sd=deriv(Out_120318_fd$seconds, Out_120318_fd$fd),
    sd11=deriv(Out_120318_fd$seconds, Out_120318_fd$fd11),
    sd25=deriv(Out_120318_fd$seconds, Out_120318_fd$fd25)
  )

#Force column names of the second derivative df 
colnames(Out_120318_sd, do.NULL = FALSE)
colnames(Out_120318_sd)<-c("seconds", "sd", "sd11", "sd25")

sd_120318_plot<-ggplot(Out_120318_sd)+
  #geom_point(aes(seconds, sd),colour = "green")+
  geom_point(aes(seconds, sd11),colour = "blue")+
  geom_point(aes(seconds, sd25), colour = "red")+
  annotation_custom(label1)+
  annotation_custom(label2)+
  annotation_custom(label3)+
  annotation_custom(label4)+
  annotation_custom(label5)+
  annotation_custom(label6)+
  annotation_custom(label7)+
  annotation_custom(label8)+
  scale_y_continuous(limits = c(-5, 5))+
  labs(x="Seconds", title="second derivative")+
  theme_classic()

print(sd_120318_plot)
ggplotly(sd_120318_plot)
```

16 - 3367 - 3668
15 - 7824 - 8125
14 - 13046 - 13347
