---
title: "November 12 2018"
author: "Rachel Havranek"
date: "12/3/2018"
output: html_document
---
# Load Libraries and functions 
```{r import libraries and functions, echo= FALSE, message=FALSE}
#setwd("/Volumes/HAVRANEK18/Lab_testing_flasks")  
library(ggplot2)
library (plotly)
library (lubridate)
library(tidyverse)
library(grid)
library(hms)
library(fpp2)
library(zoo)

#derivative function 
deriv <- function(x, y) {
  diffvar <- diff(y) / diff(x)
}

peaks.read <- function(filepath){
  	files <-  list.files(path=filepath, recursive = T,full.names = T) 
	#lists the file names in the filepath directory
  	tables <- lapply(files, read.table,header = T, stringsAsFactors = F) 
	#reads in the files
  	df <- do.call("rbind",tables) #combines all the tables into one dataframe
  	df2 <- df[,c(1:2,17:19)] #subsets the dataframe to include only necessary columns
  	df3 <- df2[complete.cases(df2),] #subsets the dataframe to include only 
	#complete cases, eliminating rows at the end that aren't complete
  	return(df3)
	} #peaks.read is a function that will read files from all sub-folders of the
	#directory named in filepath and combine them into one dataframe

h2o_slope <- function(data,window){
  H2O_Slope_Window <- numeric()
  for (i in 1:nrow(data)){    
    if(i<=window/2){                                                                   
      H2O_Slope_Window[i] = 
        as.numeric(coef(lm(data$H2O[1:(i+window/2)]~
		data$seconds[1:(i+window/2)]))[2])
    }else{
      if(i >= (nrow(data)-window/2)){
        H2O_Slope_Window[i] = 
          as.numeric(coef(lm(data$H2O[(i-window/2):nrow(data)]~
		data$seconds[(i-window/2):nrow(data)]))[2])
      }else{
        H2O_Slope_Window[i] = 
          as.numeric(coef(lm(data$H2O[(i-window/2):(i+window/2)]~
		data$seconds[(i-window/2):(i+window/2)]))[2])       
      }
    }
  }
  return(H2O_Slope_Window)
} #h2o_slope will calculate the slope of h2o vs. experiment second in a moving window  
```

# Load Data 
```{r}
# folder_111218 <- "/Volumes/HAVRANEK18/111218"
#  data_111218 <- peaks.read(folder_111218) 
#  saveRDS(data_111218, "/Volumes/HAVRANEK18/Lab_testing_flasks/data/data_111218.RDS")
data_111218 <- readRDS("/Volumes/HAVRANEK18/Lab_testing_flasks/data/data_111218.RDS")

# I did a lot of work between 5 - 10 pm so I need to bring in the data from the fifth b/c the picarro uses GMT time 

# folder_111318 <- "/Volumes/HAVRANEK18/111318"
# data_111318 <- peaks.read(folder_111318)
# saveRDS(data_111318, "/Volumes/HAVRANEK18/Lab_testing_flasks/data/data_111318.RDS")
data_111318 <- readRDS("/Volumes/HAVRANEK18/Lab_testing_flasks/data/data_111318.RDS")
```

# Add in some Metadata 
This chunk (1) creates a date/time column to match my timezone, so I can use my lab notes as a guide for data reduction, (2) adds in a linearity correction column for later comparison, (3) parses the data down to when I was running experiments, and (4) plots that water concentration data 

using the instrument specific correction of:  *This is peice isn't ready to go yet* 

Determined by altering push flow: 
\[\delta^{18}O_{corr} = \delta^{18}O_{raw}+ (-0.0002*[H2O]+4.8021)\]
\[\delta^{2}H = \delta^{2}H_{raw} + ( -0.0004*[H2O]+5.3133)\]
```{r}
data_111218 <- data_111218 %>% mutate(
  lin_d18O = (-0.0002*H2O+4.8021) + Delta_18_16,
    lin_dD = (-0.0004*H2O +5.31 ) + Delta_D_H,
  MST = 
    paste(data_111218$DATE, data_111218$TIME) %>% 
    ymd_hms() %>% 
    with_tz(tzone=c("America/Denver"))
)

data_111318 <- data_111318 %>% mutate(
  lin_d18O = (-0.0002*H2O+4.8021) + Delta_18_16,
  lin_dD = (-0.0004*H2O +5.31 ) + Delta_D_H,
  MST = 
    paste(data_111318$DATE, data_111318$TIME) %>% 
    ymd_hms() %>% 
    with_tz(tzone=c("America/Denver"))
)

evening_111218 <- data_111318 %>% filter( MST < "2018-11-12 19:00:00")  
 
#Combine data from two different files 
data_111218 <- bind_rows(data_111218, evening_111218)

#Filter to the relevant time 
data_111218 <- data_111218 %>% filter(MST > "2018-11-12 16:00:00")

all_111218 <- ggplot(data_111218, aes(MST, H2O))+
  labs(x="Time (MST)", title="November 12, 2018")+
  geom_point()+
  theme_classic()

all_111218
```

```{r}
label1 <- grobTree(textGrob("Jumper", x=0.2,  y=0.95, hjust=0,
  gp=gpar(col="red", fontsize=13, fontface="italic")))

label2 <- grobTree(textGrob("Bottle 2", x=0.35,  y=0.95, hjust=0,
  gp=gpar(col="red", fontsize=13, fontface="italic")))

label3 <- grobTree(textGrob("Jumper", x=0.5,  y=0.95, hjust=0,
  gp=gpar(col="red", fontsize=13, fontface="italic")))

fill_concen_111218 <- ggplot(data_111218, aes(MST, H2O))+
  labs(x="Time (MST)", title="November 04, 2018")+
  geom_point()+
  theme_classic()+
  annotation_custom(label1)+
  annotation_custom(label2)+
  annotation_custom(label3)

print(fill_concen_111218)

```

```{r}
first_stand <- filter (data_111218, MST > "2018-11-12 16:42:00" & MST < "2018-11-12 16:52:00" )
df1_1 <- data.frame(
  "Time averaged" = 10,
  "Type" = "Correction",
  "valco_position" = 1,
  "mean_d18O" = mean (first_stand$Delta_18_16),
  "sd_d18O" = sd (first_stand$Delta_18_16),
  "mean_dD" = mean(first_stand$Delta_D_H),
  "sd_dD" = sd(first_stand$Delta_D_H),
  "mean_lin_d18O" = mean(first_stand$lin_d18O),
  "mean _lin_dD"= mean(first_stand$lin_dD),
  "sd_lin_d18O" = sd(first_stand$lin_d18O),
  "sd _lin_dD"= sd(first_stand$lin_dD),
  stringsAsFactors = FALSE
)

fill_jumper_111218 <- filter (data_111218, MST > "2018-11-12 17:34:00" & MST < "2018-11-12 17:44:00")
df1 <- data.frame(
  "Time averaged" = 10,
  "Type" = "Correction",
  "valco_position" = 1,
  "mean_d18O" = mean (fill_jumper_111218$Delta_18_16),
  "sd_d18O" = sd (fill_jumper_111218$Delta_18_16),
  "mean_dD" = mean(fill_jumper_111218$Delta_D_H),
  "sd_dD" = sd(fill_jumper_111218$Delta_D_H),
  "mean_lin_d18O" = mean(fill_jumper_111218$lin_d18O),
  "sd_lin_d18O" = sd(fill_jumper_111218$lin_d18O),
  "mean _lin_dD"= mean(fill_jumper_111218$lin_dD),
  "sd _lin_dD"= sd(fill_jumper_111218$lin_dD),
  stringsAsFactors = FALSE
)


```

##  Pull out results data and add in time averaging 
```{r}
Out_111218 <- evening_111218 %>% 
  filter( MST >"2018-11-12 16:52:00" & MST < "2018-11-12 17:29:00")

Out_111218$seconds <- as.numeric(row.names(Out_111218))
 
Out_111218 <- Out_111218 %>% 
  mutate(
  m_11= rollmean(H2O, k=11, fill=NA),
  m_25=rollmean(H2O, k=25, fill=NA)) %>% 
  subset(seconds >13 & seconds<(nrow(Out_111218)-13))
```

## Calulate the first derivative and second derivative of water concentration data, then plot the second derivative 
```{r}
#Initialize data frame
Out_111218_fd<-as.data.frame(1:(nrow(Out_111218)-1))

# mutate in the first derivatives, and the rolling first derivatives 
Out_111218_fd<- Out_111218_fd %>% 
  mutate(
    fd = deriv(Out_111218$seconds, Out_111218$H2O),
    fd11=deriv(Out_111218$seconds, Out_111218$m_11),
    fd25=deriv(Out_111218$seconds, Out_111218$m_25)
  )

#Force column names so I can do the second derivative 
colnames(Out_111218_fd, do.NULL = FALSE)
colnames(Out_111218_fd)<-c("seconds", "fd", "fd11", "fd25")

#Initialize second derivative data frame
Out_111218_sd <- as.data.frame(1:(nrow(Out_111218_fd)-1))

#Mutate in the second derivatives 
Out_111218_sd <- Out_111218_sd %>% 
  mutate(
    sd=deriv(Out_111218_fd$seconds, Out_111218_fd$fd),
    sd11=deriv(Out_111218_fd$seconds, Out_111218_fd$fd11),
    sd25=deriv(Out_111218_fd$seconds, Out_111218_fd$fd25)
  )

#Force column names of the second derivative df 
colnames(Out_111218_sd, do.NULL = FALSE)
colnames(Out_111218_sd)<-c("seconds", "sd", "sd11", "sd25")

sd_111218_plot<- ggplot(Out_111218_sd)+
  geom_point(aes(seconds, sd),colour = "green")+
  geom_point(aes(seconds, sd11),colour = "blue")+
  geom_point(aes(seconds, sd25), colour = "red")+
  annotation_custom(label1)+
  annotation_custom(label2)+
  annotation_custom(label3)+
  geom_vline (xintercept = 200)+
  scale_y_continuous(limits = c(-5, 5))+
  labs(x="Seconds", title="second derivative")+
  theme_classic()

print(sd_111218_plot)
#ggplotly(sd_111218_plot)
```

Based on the plotly I'm going to throw away the first 200 seconds, but an argument could be made to move that line a little to the right. 

# Results and smoosh 
```{r}
results_111218_2 <- filter (Out_111218, seconds > 200 & seconds < 501)

R2_df <- data.frame(
  "valco_position" = 2,
  "Time averaged" = 5,
  "Type" = "result",
  "mean_d18O" = mean (results_111218_2$Delta_18_16),
  "sd_d18O" = sd (results_111218_2$Delta_18_16),
  "mean_dD" = mean(results_111218_2$Delta_D_H),
  "sd_dD" = sd(results_111218_2$Delta_D_H),
  "mean_lin_d18O" = mean(results_111218_2$lin_d18O),
  "sd_lin_d18O" = sd (results_111218_2$lin_d18O),
  "mean _lin_dD"= mean(results_111218_2$lin_dD),
  "sd_lin_dD" = sd (results_111218_2$lin_dD),
  stringsAsFactors = FALSE
)

summary_111218 <-  bind_rows(df1_1, R2_df, df1 )
write.csv(summary_111218, "/Volumes/HAVRANEK18/Lab_testing_flasks/data_output/111218 Summary.csv")
```



