---
title: "November_4_2018"
author: "Rachel Havranek"
date: "11/12/2018"
output: html_document
---

## Load Libraries and functions 
```{r import libraries and functions, message=FALSE}
setwd("/Volumes/HAVRANEK18")  
library(ggplot2)
library (plotly)
library (lubridate)
library(tidyverse)
library(grid)
library(hms)
library(fpp2)
library(zoo)

#derivative function 
deriv <- function(x, y) {
  diffvar <- diff(y) / diff(x)
}

peaks.read <- function(filepath){
  	files <-  list.files(path=filepath, recursive = T,full.names = T) 
	#lists the file names in the filepath directory
  	tables <- lapply(files, read.table,header = T, stringsAsFactors = F) 
	#reads in the files
  	df <- do.call("rbind",tables) #combines all the tables into one dataframe
  	df2 <- df[,c(1:2,17:19)] #subsets the dataframe to include only necessary columns
  	df3 <- df2[complete.cases(df2),] #subsets the dataframe to include only 
	#complete cases, eliminating rows at the end that aren't complete
  	return(df3)
	} #peaks.read is a function that will read files from all sub-folders of the
	#directory named in filepath and combine them into one dataframe

h2o_slope <- function(data,window){
  H2O_Slope_Window <- numeric()
  for (i in 1:nrow(data)){    
    if(i<=window/2){                                                                   
      H2O_Slope_Window[i] = 
        as.numeric(coef(lm(data$H2O[1:(i+window/2)]~
		data$seconds[1:(i+window/2)]))[2])
    }else{
      if(i >= (nrow(data)-window/2)){
        H2O_Slope_Window[i] = 
          as.numeric(coef(lm(data$H2O[(i-window/2):nrow(data)]~
		data$seconds[(i-window/2):nrow(data)]))[2])
      }else{
        H2O_Slope_Window[i] = 
          as.numeric(coef(lm(data$H2O[(i-window/2):(i+window/2)]~
		data$seconds[(i-window/2):(i+window/2)]))[2])       
      }
    }
  }
  return(H2O_Slope_Window)
} #h2o_slope will calculate the slope of h2o vs. experiment second in a moving window  
```

## Load Data
```{r}
# folder_110418 <- "/Volumes/HAVRANEK18/110418"
# data_110418 <- peaks.read(folder_110418) 
# saveRDS(data_110418, "data_110418.RDS")
data_110418 <- readRDS("data_110418.RDS")

# I did a lot of work between 5 - 10 pm so I need to bring in the data from the fifth b/c the picarro uses GMT time 

# folder_110518 <- "/Volumes/HAVRANEK18/110518"
# data_110518 <- peaks.read(folder_110518) 
# saveRDS(data_110518, "data_110518.RDS")
data_110518 <- readRDS("data_110518.RDS")
```

## 1. Use Lubridate to match Picarro time with lab Notes
## 2. Parse the data down to when I was running experiments, and plot that water concentration data 

```{r}
data_110418 <- data_110418 %>% mutate(
  MST = 
    paste(data_110418$DATE, data_110418$TIME) %>% 
    ymd_hms() %>% 
    with_tz(tzone=c("America/Denver"))
)

data_110518 <- data_110518 %>% mutate(
  MST = 
    paste(data_110518$DATE, data_110518$TIME) %>% 
    ymd_hms() %>% 
    with_tz(tzone=c("America/Denver"))
)

evening_110418 <- data_110518 %>% 
  filter( MST < "2018-11-04 23:00:00")

data_110418 <- bind_rows(data_110418, evening_110418)

all_110418 <- ggplot(data_110418, aes(MST, H2O))+
  labs(x="Time (MST)", title="November 04, 2018")+
  geom_point()+
  theme_classic()

all_110418
```

## Plot the filling stage with grob labels 
```{r}
fill_110418 <- evening_110418 %>% 
  filter( MST < "2018-11-04 20:00:00")

#These labels help me visually in my plots 
label1 <- grobTree(textGrob("Bottle 16", x=0.1,  y=0.95, hjust=0,
  gp=gpar(col="red", fontsize=13, fontface="italic")))

label2 <- grobTree(textGrob("Bottle 15", x=0.35,  y=0.95, hjust=0,
  gp=gpar(col="red", fontsize=13, fontface="italic")))

label3 <- grobTree(textGrob("Bottle 14", x=0.55,  y=0.95, hjust=0,
  gp=gpar(col="red", fontsize=13, fontface="italic")))

label4 <- grobTree(textGrob("Bottle 2", x=0.8,  y=0.95, hjust=0,
  gp=gpar(col="red", fontsize=13, fontface="italic")))

label5 <- grobTree(textGrob("Jumper", x=0.9,  y=0.95, hjust=0,
  gp=gpar(col="red", fontsize=13, fontface="italic")))

fill_concen_110418 <- ggplot(fill_110418, aes(MST, H2O))+
  labs(x="Time (MST)", title="November 03, 2018")+
  geom_point()+
  theme_classic()+
  annotation_custom(label1)+
  annotation_custom(label2)+
  annotation_custom(label3)+
  annotation_custom(label4)+
  annotation_custom(label5)

print(fill_concen_110418)
```


## Fill flasks with water vapor, and reduce the input values 
```{r}
fill_16_110418 <- filter(evening_110418, MST > "2018-11-04 17:27:00" & MST < "2018-11-04 17:37:00")
df16 <- data.frame(
  "valco_position" = 16,
  "mean_d18O" = mean (fill_16_110418$Delta_18_16),
  "sd_d18O" = sd (fill_16_110418$Delta_18_16),
  "mean_dD" = mean(fill_16_110418$Delta_D_H),
  "sd_dD" = sd(fill_16_110418$Delta_D_H),
  stringsAsFactors = FALSE
)

fill_15_110418 <- filter (evening_110418, MST > "2018-11-04 18:08:00" & MST < "2018-11-04 18:18:00")
df15 <- data.frame(
  "valco_position" = 15,
  "mean_d18O" = mean (fill_15_110418$Delta_18_16),
  "sd_d18O" = sd (fill_15_110418$Delta_18_16),
  "mean_dD" = mean(fill_15_110418$Delta_D_H),
  "sd_dD" = sd(fill_15_110418$Delta_D_H),
  stringsAsFactors = FALSE
)

fill_14_110418 <- filter (evening_110418, MST > "2018-11-04 18:48:00" & MST < "2018-11-04 18:58:00")
df14 <- data.frame(
  "valco_position" = 14,
  "mean_d18O" = mean (fill_14_110418$Delta_18_16),
  "sd_d18O" = sd (fill_14_110418$Delta_18_16),
  "mean_dD" = mean(fill_14_110418$Delta_D_H),
  "sd_dD" = sd(fill_14_110418$Delta_D_H),
  stringsAsFactors = FALSE
)

fill_2_110418 <- filter (evening_110418, MST > "2018-11-04 19:37:00" & MST < "2018-11-04 19:47:00")
df2 <- data.frame(
  "valco_position" = 2,
  "mean_d18O" = mean (fill_2_110418$Delta_18_16),
  "sd_d18O" = sd (fill_2_110418$Delta_18_16),
  "mean_dD" = mean(fill_2_110418$Delta_D_H),
  "sd_dD" = sd(fill_2_110418$Delta_D_H),
  stringsAsFactors = FALSE
)

fill_jumper_110418 <- filter (evening_110418, MST > "2018-11-04 19:50:00" & MST < "2018-11-04 19:59:00")
df1 <- data.frame(
  "valco_position" = 1,
  "mean_d18O" = mean (fill_jumper_110418$Delta_18_16),
  "sd_d18O" = sd (fill_jumper_110418$Delta_18_16),
  "mean_dD" = mean(fill_jumper_110418$Delta_D_H),
  "sd_dD" = sd(fill_jumper_110418$Delta_D_H),
  stringsAsFactors = FALSE
)

fill_110418_summary <-  bind_rows(df16, df15, df14, df2, df1)
```

##  Pull out results data and add in time averaging 
```{r}
Out_110418 <- evening_110418 %>% 
  filter( MST >"2018-11-04 20:00:00" & MST < "2018-11-04 22:00:00")

Out_110418$seconds <- as.numeric(row.names(Out_110418))
 
Out_110418 <- Out_110418 %>% 
  mutate(
  m_11= rollmean(H2O, k=11, fill=NA),
  m_25=rollmean(H2O, k=25, fill=NA)) %>% 
  subset(seconds >13 & seconds<(nrow(Out_110418)-13))
```

## Calulate the first derivative and second derivative of water concentration data, then plot the second derivative 
```{r}
#Initialize data frame
Out_110418_fd<-as.data.frame(1:(nrow(Out_110418)-1))

# mutate in the first derivatives, and the rolling first derivatives 
Out_110418_fd<- Out_110418_fd %>% 
  mutate(
    fd = deriv(Out_110418$seconds, Out_110418$H2O),
    fd11=deriv(Out_110418$seconds, Out_110418$m_11),
    fd25=deriv(Out_110418$seconds, Out_110418$m_25)
  )

#Force column names so I can do the second derivative 
colnames(Out_110418_fd, do.NULL = FALSE)
colnames(Out_110418_fd)<-c("seconds", "fd", "fd11", "fd25")

#Initialize second derivative data frame
Out_110418_sd <- as.data.frame(1:(nrow(Out_110418_fd)-1))

#Mutate in the second derivatives 
Out_110418_sd <- Out_110418_sd %>% 
  mutate(
    sd=deriv(Out_110418_fd$seconds, Out_110418_fd$fd),
    sd11=deriv(Out_110418_fd$seconds, Out_110418_fd$fd11),
    sd25=deriv(Out_110418_fd$seconds, Out_110418_fd$fd25)
  )

#Force column names of the second derivative df 
colnames(Out_110418_sd, do.NULL = FALSE)
colnames(Out_110418_sd)<-c("seconds", "sd", "sd11", "sd25")

sd_110418_plot<-ggplot(Out_110418_sd)+
  geom_point(aes(seconds, sd),colour = "green")+
  geom_point(aes(seconds, sd11),colour = "blue")+
  geom_point(aes(seconds, sd25), colour = "red")+
  annotation_custom(label1)+
  annotation_custom(label2)+
  annotation_custom(label3)+
  annotation_custom(label4)+
  scale_y_continuous(limits = c(-5, 5))+
  labs(x="Seconds", title="second derivative")+
  theme_classic()

print(sd_110418_plot)
ggplotly(sd_110418_plot)
```
From this chunk, determine the seconds (in the Out_110418 data frome) where the second derivative of water concentration is approximately 0. Use the first 300 seconds (5 minutes) where the system is draining at steady state. Then feed that into the data frames in the next chunk 

## Calculate results using times when the second derivative of water concentration ~0
```{r}
#Create a data frame for each bottle 
results_110418_16 <- filter(Out_110418, seconds > 900 & seconds < 1200)
results_110418_15 <- filter (Out_110418, seconds > 2750 & seconds < 3050)
results_110418_14 <- filter (Out_110418, seconds > 4600 & seconds < 4900)
results_110418_2 <- filter (Out_110418, seconds > 6500 & seconds < 6800)
results_110418_1 <- filter (Out_110418, seconds > 7874 & seconds < 8400)

#Create a data frame for each bottle and then smoosh the dfs together
R16_df <- data.frame(
  "valco_position" = 16,
  "Time averaged" = 5,
  "Type" = "result",
  "mean_d18O" = mean (results_110418_16$Delta_18_16),
  "sd_d18O" = sd (results_110418_16$Delta_18_16),
  "mean_dD" = mean(results_110418_16$Delta_D_H),
  "sd_dD" = sd(results_110418_16$Delta_D_H),
  stringsAsFactors = FALSE
)

R15_df <- data.frame(
  "valco_position" = 15,
  "Time averaged" = 5,
  "Type" = "result",
  "mean_d18O" = mean (results_110418_15$Delta_18_16),
  "sd_d18O" = sd (results_110418_15$Delta_18_16),
  "mean_dD" = mean(results_110418_15$Delta_D_H),
  "sd_dD" = sd(results_110418_15$Delta_D_H),
  stringsAsFactors = FALSE
)

R14_df <- data.frame(
  "valco_position" = 14,
  "Time averaged" = 5,
  "Type" = "result",
  "mean_d18O" = mean (results_110418_14$Delta_18_16),
  "sd_d18O" = sd (results_110418_14$Delta_18_16),
  "mean_dD" = mean(results_110418_14$Delta_D_H),
  "sd_dD" = sd(results_110418_14$Delta_D_H),
  stringsAsFactors = FALSE
)

R2_df <- data.frame(
  "valco_position" = 2,
  "Time averaged" = 5,
  "Type" = "result",
  "mean_d18O" = mean (results_110418_2$Delta_18_16),
  "sd_d18O" = sd (results_110418_2$Delta_18_16),
  "mean_dD" = mean(results_110418_2$Delta_D_H),
  "sd_dD" = sd(results_110418_2$Delta_D_H),
  stringsAsFactors = FALSE
)

R1_df<- data.frame(
  "valco_position" = 1,
  "Time averaged" = 8.76,
  "Type" = "Correction",
  "mean_d18O" = mean (results_110418_1$Delta_18_16),
  "sd_d18O" = sd (results_110418_1$Delta_18_16),
  "mean_dD" = mean(results_110418_1 $Delta_D_H),
  "sd_dD" = sd(results_110418_1 $Delta_D_H),
  stringsAsFactors = FALSE
)

results_110418_summary <- bind_rows(R16_df, R15_df, R14_df, R2_df, R1_df)
```

## Bind together input and results, write to csv 
```{r}
Summary_110418 <- bind_rows(fill_110418_summary, results_110418_summary)
print(Summary_110418)
#write.csv(Summary_110418, "November 04, 2018 Summary table.csv")
```


